---
title: "monarchMakeRDS"
author: "Jerad Acosta"
date: "May 14, 2015"
output: html_document
---

```{r}
# Set working Directory to CSV data location
## changed file to data location
wd <- "~/Documents/Data-Apps/monarch-dashboard"
setwd(wd)

# load original data set into R for Manipulation
filePath <- "./Monarch_Original.csv"
data <- read.csv(filePath)

# Set working directory to Shinny app
## Just create RDS the save to ./data directory

# Create data Directory for .rds file after done manipulating it
if(!file.exists("./monarch-app/data")){dir.create("./monarch-app/data")}
```


### Newly edited data does not need this step
```{r, reduce empty factors}
# find and remove empty factor level

# r <- which(data$timestamp == "")
# data[r,]
# data <- data[-c(r),]
```

## ReName Variables
```{r, rename-variables}
# Done by hand in Numbers / Excel and Saved in Directory
```


## Transcibing Variables

### Correct Date-Time
Current CSV file has var = "data_and_time_of_misbehavior" in class Factor
Change to class "POSIXlt" "POSIXt"
```{r, create POSIX date-time var}
data$date_time <- strptime(as.character(data$date_and_time_of_misbehavior), format = "%Y-%m-%d %H:%M:%S")
# Create Numerical Time for Easier Graphical Computing
```

### Create $environment variable then cast
```{r}
# Create environment factor variable
# Search for word class to create a classroom level in the environmental factors
# Note for personal learning: Which ever factor level is assigned first is the one that stays
# Just like an If / Else statement - Priority goes to first statement
data$environment <- ifelse(grepl("class", data$location_of_misbehavior, ignore.case = TRUE), "Classroom",
                           ifelse(grepl("PE", data$location_of_misbehavior, ignore.case = TRUE), "PE", 
                                  ifelse(grepl("lunch", data$location_of_misbehavior, ignore.case = TRUE), "Lunch",
                                         ifelse(grepl("yoga", data$location_of_misbehavior, ignore.case = TRUE), "Yoga",
                                                ifelse(grepl("gym", data$location_of_misbehavior, ignore.case = TRUE), "Gym", 
                                                       ifelse(grepl("blacktop", data$location_of_misbehavior, ignore.case = TRUE), "Blacktop",
                                                              ifelse(grepl("cafeteria", data$location_of_misbehavior, ignore.case = TRUE), "Cafeteria",
                                                                     ifelse(grepl("hallway", data$location_of_misbehavior, ignore.case = TRUE), "Hallway",
                                                                            ifelse(grepl("playground", data$location_of_misbehavior, ignore.case = TRUE), "Playground",
                                                                                   ifelse(grepl("snack", data$location_of_misbehavior, ignore.case = TRUE), "Snack Time", NA))))))))))
```

#### cast $environment variable class factor (currently character)
**TODO:** 
[X] This results in 45 levels which should be reduced to necessary and _other_
adding NA as the else statement fixed this

```{r, environment-var-to-factor}
data$environment <- as.factor(data$environment)
```

## Reduce Number of Variables

To reduce the size of the data object we should remove variables we are not going to be using in the visualizations for this application

### Remove Replaced Variables
+ date_and_time_of_misbehavior
+ location_of_misbehavior
+ time_stamp

#### DID NOT DO
Saved variables since RDS file is still small and the project is still in development
```{r}
drop <- c("date_and_time_of_misbehavior", "location_of_misbehavior", "time_stamp")
data <- data[,!(names(data) %in% drop)]
```

### Reduce duplicated Variables
+ documenting_staff_id & documenting_staff
+ reporting_staff_id & reporting_staff

### Variables not being used (yet)
+ classroom_or_administrative_managed
+ narrative_discription_of_misbehavior
    + could use NLP algorithm for future use
+ d12_planning_completed
+ narrative_of_concequence


### Need to Transcribe
+ type_of_misbehavior
+ concequence


## Write RDS file
using saveRDS method to write and readRDS to load
```{r, save-rds}
saveRDS(data, file = "./monarch-app/data/student-referrals.rds")
```